---
title: "Avocado Satiety Study"
output: github_document
always_allow_html: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)

# Setup -------------------------------------------------------------------

Sys.setlocale("LC_CTYPE", "English_United States.utf8")

# Required packages
pacs <- c(
  "tidyverse", 
  "readxl", 
  "janitor", 
  "hms", 
  "tableone", 
  "kableExtra",
  "ggridges", 
  "patchwork",
  "nlme",
  "emmeans"
)

sapply(pacs, require, character.only = TRUE)
source("functions.R")

```

## Description

* A secondary analysis of the Habitual Diet and Avocado Trial (HAT), a randomized clinical trial of two-arm parallel design
* This study examines how avocado consumption affects energy intake, energy density, and meal intervals
* The unit of analysis is eating occasion, either meals or snacks

## Datasets

* There are five data files:
  * `macro_protocol3.xlsx`
    * Includes patient demographics
    * n = 900 participants, `cpartid` are all distinct
    
  * `HAT_SCMeal.csv`
    * Meal-level data from NDSR
    * Used to identify meals containing avocado (`rfgscmFRU500`, in serving size)
    * n = 19,643 meals
    
  * `meal_timdiff_updated.xlsx`
    * Also meal-level data, but time of the meal and meal order were manually corrected
    * Includes `Coding` -- what types of error was identified for each meal, if any  
    * n = 19,643 meals
    
  * `Gram_kcal_new2.csv`
    * Also meal-level data, but includes total intake as well as food intake (excluding beverages)
    * n = 19,643 meals
    
  * `HAT_Intake.csv`
    * recall-day-level data
    * Includes day of the week, total kcal per day
    * n = 3901 recall days

* Five datasets were merged together to create an analytic dataset

```{r data_prep}

# Read data ---------------------------------------------------------------

# The following 5 files were used to create an analytic data
# Subject ID is: pid  (convert cpartid to pid if necessary)

# Protocol file
# Includes all participants who had two or more compliant recalls
# Will be used as demographic data
# n = 900 (cpartid all distinct)
protocol <- read_excel("./data/macro_protocol3.xlsx") %>%
  clean_names() %>%
  rename(pid = cpartid) %>%
  select(pid, site_id, randarm, education, sex, age, bmi, waist, race)

# Contains food group serving counts at the meal or eating occasion level (FGSCMeal),
# Used to identify meals with avocado
# Avocado intake = rfgscmFRU500
# n = 19,643
meal_servings <- read_csv("./data/HAT_SCMeal.csv") %>%
  mutate(tmtime = as.numeric(tmtime))        # readxl reads times as hms; convert to seconds

# Revised Meal file (per meal)
# Contains corrected meal time 
# n = 19,643
meals_updated <- read_excel("./data/meal_timdiff_updated.xlsx") %>%
  clean_names() %>%
  rename(pid = cpartid) %>%
  mutate(
    tmtime = as_hms(tmtime),                 # tmtime read as datetime, convert to hms
    tmtime = as.numeric(tmtime))             # convert to seconds

# Total intake (gram, kcal) per meal, separating beverages 
# Appears to be based on meal file
# This will be used instead of meal file
# n = 19,643
non_beverages <- read_csv("./data/Gram_kcal_new2.csv") %>%
  mutate(pid = as.numeric(cpartid)) %>%
  select(-cpartid)

# Intake (per day) file
# Contains total nutrient intake per day for each participant
# Needed: intake day of the week
# idow: 0 = Sunday, 1 = Monday,..., 6 = Saturday 
# n = 3,901
intake <- read_csv("./data/HAT_Intake.csv") %>%
  mutate(weekday = if_else(idow %in% c(0, 6), 0L, 1L),
         kcal_per_day = rikcal) %>%
  select(pid, visitcode, weekday, kcal_per_day, idow)


# Identify meals with avocado ---------------------------------------------

# Assume one whole avocado = 168 gram
# One whole avocado = 2.24 serving size

# Avocado intake categories
avocado_intake_cat_labels <- c(
  "No avocado at all", 
  "<1 avocado", 
  "1 whole avocado or more" 
)

# Based on meal_serving file
# Order by pid, visit, and meal sequence
# Identify meals with avocado intake
meal_servings2 <- meal_servings %>%
  arrange(pid, visitcode, cmealid) %>%
  group_by(pid, visitcode) %>%
  mutate(
    meal_with_avocado  = if_else(rfgscmFRU0500 > 2.23, 1L, 0L),
    avocado_gram       = rfgscmFRU0500 * 168 / 2.24,
    avocado_intake_cat = case_when(
      avocado_gram == 0                        ~ 1L,
      avocado_gram > 0   & avocado_gram < 168  ~ 2L,
      avocado_gram >= 168                      ~ 3L,
      
    ),
    avocado_intake_cat = factor(avocado_intake_cat, levels = 1:3, labels = avocado_intake_cat_labels),
    # meal_after_avocado = if_else(cmealid == 1, NA_integer_, lag(meal_with_avocado)),
    # previous_avcd_intk = if_else(cmealid == 1, NA_real_,    lag(rfgscmFRU0500)),
    # previous_meal      = if_else(cmealid == 1, NA_real_,    lag(as.numeric(imname)))
    # tmtime             = if_else(tmtime == 0 , 3600 * 24,   as.numeric(tmtime)),
    # meal_interval      = if_else(cmealid == 1, NA_real_,    (tmtime - lag(tmtime)) / 3600)
  ) %>%
  ungroup() %>%
  select(
    pid, 
    visitcode, 
    cmealid, 
    imname, 
    rfgscmFRU0500,
    meal_with_avocado,
    avocado_gram,
    avocado_intake_cat,
    # meal_after_avocado, 
    # previous_avcd_intk, 
    # previous_meal 
    # meal_interval
  )

# analytic_df %>% 
  # filter(is.na(meal_with_avocado))

# Correcting meal names ---------------------------------------------------

# Based on meals_updated
# meal_freq will be re-calculated at later step
meals_updated2 <- meals_updated %>%
  arrange(pid, visitcode, cmealid) %>%
  group_by(pid, visitcode) %>%
  mutate(
    tmtime        = if_else(tmtime == 0 & cmealid > 1, 3600 * 24,   as.numeric(tmtime)),
    meal_interval = if_else(cmealid == 1 | coding %in% c(3, 5), NA_real_, (tmtime - lag(tmtime)) / 3600),
    coding        = if_else(coding == 0, 2L, as.integer(coding)),
    # meal_freq     = sum(imname != 8)
  ) %>%
  ungroup() %>%
  rename(tmtime_fixed = tmtime) %>%
  select(pid, visitcode, cmealid, imname, tmtime_fixed, meal_interval, coding)


# Merge files for analytic dataset ----------------------------------------

# Time threshold (in seconds)
t11 <- 11 * 3600
t12 <- 12 * 3600
t17 <- 17 * 3600 

# Category labels
race_labels      <- c("African American", "Asian", "Caucasian", "Other")
educ_labels      <- c("None", "Vocational school", "College degree", "Graduate degree")
meal_type_labels <- c(
  "Morning meal", 
  "Morning snack", 
  "Midday meal", 
  "Midday snack", 
  "Evening meal", 
  "Late night snack",
  "Beverage"
)
meal_place_labels <- c(
  "Home/Friend's home",
  "Work/School",
  "Deli/Take-out/Restaurant",
  "Daycare/Party/Travelling/Other"
)

# Merge files
# Begin with meals_updated2
analytic_df <- meals_updated2 %>%

  # Inner-join non_beverages
  # Calculate total density, food kcal/grams, and food density
  # Note that food density > 9 is set to missing (5 meals affected)
  inner_join(
    non_beverages %>%
    select(-imname) %>% 
      mutate(
        total_density = total_kcal / total_gram,
        food_kcal     = total_kcal - bvrg_kcal,
        food_gram     = total_gram - bvrg_gram,
        food_density  = food_kcal / food_gram,
        food_density  = if_else(food_density > 9, NA_real_, food_density)
      ),
    by = c("pid", "visitcode", "cmealid")
  ) %>%
  
  # Inner-join meal_servings2 
  inner_join(meal_servings2 %>% select(-imname), by = c("pid", "visitcode", "cmealid")) %>%
  
  # Restrict to per-protocol participants; add demographics
  inner_join(protocol, by = "pid") %>%
  
  # Inner-join with intake per day file: add day_of_wk, kcal_per_day 
  inner_join(intake,   by = c("pid", "visitcode")) %>%
  
  # Correct 3 meals miscoded as beverages
  mutate(imname = case_when(
    pid == 15120643 & visitcode == "DT4" & cmealid == 1 ~ 4L,
    pid == 15131688 & visitcode == "DT4" & cmealid == 1 ~ 1L,
    pid == 10004467 & visitcode == "DT2" & cmealid == 7 ~ 4L,
    TRUE ~ as.integer(imname)
  )) %>%

  # Recode variables 
  mutate(
    bmi_cat = case_when(
      bmi < 25             ~ 1,
      bmi >= 25 & bmi < 30 ~ 2,
      bmi >= 30            ~ 3
    ),
    bmi_cat = factor(bmi_cat, labels = c("Normal", "Overweight", "Obese")),
    
    region = case_when(
      site_id %in% c(401, 402, 501) ~ 1L,  # East Coast
      site_id %in% c(201, 301)      ~ 2L   # West Coast
      ),
    region = factor(region, labels = c("East coast", "West coast")),
    
    hour = tmtime_fixed / 3600,
    
    meal_type  = case_when(
      imname == 8                                             ~ 9L,  # Beverage
      imname == 4 & tmtime_fixed <= t12                       ~ 2L,  # Morning snack
      imname == 4 & tmtime_fixed >  t12 & tmtime_fixed <= t17 ~ 4L,  # PM snack
      imname == 4 & tmtime_fixed >  t17                       ~ 6L,  # HS snack
      tmtime_fixed <= t11                                     ~ 1L,  # Morning meal
      tmtime_fixed >  t11 & tmtime_fixed < t17                ~ 3L,  # Midday meal
      tmtime_fixed >= t17                                     ~ 5L   # Evening meal
    ),
    meal_or_snack = if_else(meal_type %in% c(2, 4, 6), 1, 0),
    meal_type     = factor(meal_type, labels = meal_type_labels),
    meal_or_snack = factor(meal_or_snack, labels = c("Meal", "Snack")), 
    
    idow = factor(idow, labels = c("Sun", "Mon", "Tue", "Wed", "Thur", "Fri", "Sat")),
    
    meal_place  = case_when(
      implace %in% c(1, 10)             ~ 1L,  # Home / Friend's home
      implace %in% c(2, 3)              ~ 2L,  # Work / School
      implace %in% c(6, 7)              ~ 3L,  # Deli/take-out / Restaurant
      implace %in% c(4, 11, 12, 13, 14) ~ 4L   # Daycare / Party / Other / Travelling
    ),
    meal_place = factor(meal_place, labels = meal_place_labels),
    
    race = factor(race, labels = race_labels),
    race = relevel(race, ref = "Caucasian"),
    
    educ_level = case_when(
      education %in% c(0, 1) ~ 1L,     # No high school / none
      education %in% c(2, 6) ~ 2L,     # Vocational / professional school
      education %in% c(3, 4) ~ 3L,     # Community / 4-year college
      education == 5         ~ 4L      # Graduate school
    ),
    educ_level = factor(educ_level, labels = educ_labels),
    
    # previous_meal_type = case_when(
    #   previous_meal %in% c(1, 2, 3, 5, 6)  ~ 0L,  # Any meal
    #   previous_meal == 4                   ~ 1L,  # Snack
    #   previous_meal == 8                   ~ 2L   # Beverage
    # ),
  ) %>%
  
  # Final exclusions: avocado arm only, no baseline, no beverages, no trace meals (total kcal >= 20, food kcal > 0.3)
  filter(
    randarm   == 1,
    visitcode != "DT1",
    total_kcal >= 20,
    imname    != 8,
    food_kcal > 0.3
  ) %>% 
  
  # Recalculate meal frequency
  add_count(pid, visitcode, name = "meal_freq") %>%
  
  # Recalculate meal interval: meal_interval_hr
  arrange(pid, visitcode, cmealid) %>% 
  group_by(pid, visitcode) %>% 
  mutate(
    meal_interval_hr = (tmtime_fixed - lag(tmtime_fixed)) / 60 ^ 2,
    meal_interval_hr = ifelse(coding %in% c(3, 5), NA_real_, meal_interval_hr),
    lag_avocado_intake_cat = lag(avocado_intake_cat)
  ) %>% 
  ungroup()

```

## Inclusion/exclusion criteria

* Note that the data has a three-level nested structure: Participants - Recall day - Meals of the day

* At the participant level:
  * Includes only those in the avocado arm

* At the recall-day level:
  * Excludes baseline food recalls (`DT1`)

* At the meal level:
  * Excludes beverage-only eating occasions
  * Excludes meals with total kcal less than 20 kcal
  * Excludes meals with food kcal less than 0.3 kcal (misclassified as meals/snacks, rather than beverages)

* After applying inclusion/exclusion criteria, the analytic data include:
  * n = 436 participants, all from the avocado arm
  * n = 1,296 recall days
    * Most of the participants (97.2%) have 3 recall days
    * Twelve participants have only 2 recall days
  * n = 4990 meals

## Outcomes

* Total energy content (kcal) of eating occasions

* Energy density (kcal / grams) of eating occasions
  * If the energy density is greater than 9, this was set to as missing (implausible)
    * These meals with density > 9 will be excluded from the analysis on energy density

* Meal interval (in hours), or time passed since the previous eating occasion of the day, excluding beverages
  * For the first eating occasion of the day, meal interval is NOT calculated (missing)
  * Some meal times were incorrectly recorded and the correct times could not be determined -- set to missing
    * (`Coding = 3 or 5`; beverage-related? See the email from AC on 5/18/2026)
  * These meals will be excluded from analysis on meal intervals


## Descriptive analysis 

### Participant characteristics

```{r participant_characteristics}

# Descriptive table -------------------------------------------------------

## Demographics -----------------------------------------------------------

demog <- analytic_df %>%
  distinct(pid, sex, race, bmi_cat, bmi, educ_level, region, waist, age)

demog_renamed <- demog %>%
  rename(
    "Age, years"              = age,
    "Sex"                     = sex,
    "Race/Ethnicity"          = race,
    "Education level"         = educ_level,
    "Region"                  = region,
    "BMI category"            = bmi_cat,
    "BMI, kg/m²"              = bmi,
    "Waist circumference, cm" = waist
  )

table_vars <- c(
  "Age, years", 
  "Sex", 
  "Race/Ethnicity", 
  "Education level",
  "Region", 
  "BMI category", 
  "BMI, kg/m²", 
  "Waist circumference, cm"
)

CreateTableOne(vars = table_vars, data = demog_renamed) %>%
  print(showAllLevels = TRUE, contDigits = 1) %>%
  kableone() %>%
  column_spec(1, bold = TRUE)

```

### Recall day

* Out of 1,296 recall days, about 30% were during weekends
  * The number of recalls on Fridays was notably lower compared to other days of the week
* The number of meals/snacks eaten across days of the week appears to be consistent
* Meal kcal per day appears to be higher on Saturdays

```{r recall_day_desc}

analytic_df %>%
  group_by(pid, visitcode, idow, kcal_per_day) %>%
  summarise(n_meals = n(), .groups = "drop") %>%
  group_by(idow) %>%
  summarise(
    n_visits = n(),
    mean_n_meals = mean(n_meals),
    mean_kcal_per_day = mean(kcal_per_day)
  ) %>%
  mutate(percent = n_visits / sum(n_visits) * 100) %>% 
  select(idow, n_visits, percent, mean_n_meals, mean_kcal_per_day) %>% 
  knitr::kable(digits = 2, col.names = c("Day of the week", "No. recalls", "%", "Mean # of meals", "Mean kcal/day"))

```

### Meal-level 

#### Meal type

* Definition: Meal type (meal period):
  * Morning meal (00:00 to 11:00)
  * Midday meal  (11:01 to 16:59) 
  * Evening meal (17:00 to 23:59)
  * Morning snack    (00:00 to 11:59) 
  * Evening snacks   (12:00 to 16:59) 
  * Late-night snack (17:00 to 23:59) 

* Frequency table by meal/snack and mean energy per meal
  * The number of morning, midday, evening meals were approximately equal
  * The number of snacks increases as the day progresses
  * The energy content of meals and snacks also increases as the day progresses

```{r meal_desc}

# Descriptive stats by meal type
tab <- analytic_df %>% 
  mutate(meal_type = droplevels(meal_type)) %>% 
  CreateTableOne(
    vars = c("total_kcal", "food_kcal"),
    strata = "meal_type",
    data = .,
    test = FALSE
  )

tab_df <- as.data.frame(print(tab, printToggle = FALSE, digits = 1)) %>%
  rownames_to_column("variable") %>%
  pivot_longer(-variable, names_to = "meal_type", values_to = "value") %>%
  pivot_wider(names_from = variable, values_from = value) %>%
  left_join(
    analytic_df %>% distinct(meal_type, meal_or_snack),
    by = "meal_type"
  ) %>%
  mutate(percent = sprintf("%.1f", parse_number(n) / nrow(analytic_df) * 100)) %>% 
  arrange(meal_or_snack) %>% 
  select(meal_type, n, percent, `total_kcal (mean (SD))`, `food_kcal (mean (SD))`)

tab_df %>% 
  knitr::kable(col.names = c("Meal/Snack", "n", "%", "Mean total kcal (SD)", "Mean food kcal (SD)"))

```

* A density plot of food kcal by meal and snack type shows a similar trend

```{r meal_density}

# Density plot of food kcal by meal/snack type
analytic_df %>% 
  filter(food_kcal > 0) %>% 
  ggplot(aes(x = food_kcal, y = fct_rev(meal_type), fill = meal_type)) +
  geom_density_ridges(alpha = 0.5) +
  theme_ridges() + 
  theme(
    legend.position = "none",
    axis.title.x = element_text(hjust = 0.5)
  ) +
  labs(x = "Food kcal per meal (log axis)", y = "") +
  scale_x_log10() +
  coord_cartesian(xlim = c(10, 5000)) +
  facet_grid(meal_or_snack ~., scales = "free_y") 

```

#### Meal place

* Meal place 
  * Among 4,990 meals, 68% of them were eaten at home or friend's home
  * The energy content was much higher when eaten at delis/restaurants or ordering takeout
  
```{r meal_place}

# Meal place and mean energy
tab <- analytic_df %>% 
  CreateTableOne(
    vars = c("total_kcal", "food_kcal"),
    strata = "meal_place",
    data = .,
    test = FALSE
  )

tab_df <- as.data.frame(print(tab, printToggle = FALSE, digits = 1)) %>%
  rownames_to_column("variable") %>%
  pivot_longer(-variable, names_to = "meal_place", values_to = "value") %>%
  pivot_wider(names_from = variable, values_from = value) %>%
  mutate(percent = sprintf("%.1f", parse_number(n) / nrow(analytic_df) * 100)) %>% 
  select(meal_place, n, percent, `total_kcal (mean (SD))`, `food_kcal (mean (SD))`)

tab_df %>% 
  knitr::kable(col.names = c("Meal place", "n", "%", "Mean total kcal (SD)", "Mean food kcal (SD)"))

```

### Outcomes

#### Energy content (food kcal)

* The distribution of food kcal per meal/snack is right-skewed 
  * The mean (SD) food kcal was 482 (355.3) kcal

```{r distr_food_kcal, fig.height = 4, fig.width = 8}

# Kcal from food meal, excluding beverages
analytic_df %>% 
  summarize(
    min     = min(food_kcal),
    Q1      = quantile(food_kcal, prob = 0.25),
    Median  = quantile(food_kcal, prob = 0.50),
    Mean    = mean(food_kcal, na.rm = TRUE),
    SD      = sd(food_kcal, na.rm = TRUE),
    Q3      = quantile(food_kcal, prob = 0.75),
    max     = max(food_kcal)
  ) %>% 
  knitr::kable(col.names = c("Min", "Q1", "Median", "Mean", "SD", "Q3", "Max"), digits = 1)

# Distribution of food kcal
# Histogram
p1 <- analytic_df %>% 
  ggplot(aes(x = food_kcal)) +
  geom_histogram() +
  labs(
    y = "Number of meals", 
    x = "Energy content per meal",
    title = "Histogram of food kcal"
  )

# ECDF
p2 <- analytic_df %>% 
  ggplot(aes(x = food_kcal)) +
  stat_ecdf() +
  scale_y_continuous(labels = scales::percent) +
  labs(
    y = "Cumlative percentage", 
    x = "Energy content per meal",
    title = "Cumulative distribution of food kcal"
  )

p1 + p2

```

#### Energy density

* The distribution of food density is slightly right-skewed 
  * The mean (SD) food density was 2.17 (1.31) kcal/g

```{r distr_food_density, fig.height = 4, fig.width = 8}

# Food density, descriptive stats 
analytic_df %>% 
  filter(!is.na(food_density)) %>% 
  summarize(
    min     = min(food_density),
    Q1      = quantile(food_density, prob = 0.25),
    Median  = quantile(food_density, prob = 0.50),
    Mean    = mean(food_density),
    SD      = sd(food_density),
    Q3      = quantile(food_density, prob = 0.75),
    max     = max(food_density)
  ) %>% 
  knitr::kable(col.names = c("Min", "Q1", "Median", "Mean", "SD", "Q3", "Max"), digits = 2)

# Distribution of food density 
# Histogram
p1 <- analytic_df %>% 
  ggplot(aes(x = food_density)) +
  geom_histogram() +
  labs(
    y = "Number of meals", 
    x = "Food density",
    title = "Histogram of food density"
  )

# ECDF
p2 <- analytic_df %>% 
  ggplot(aes(x = food_density)) +
  stat_ecdf() +
  scale_y_continuous(labels = scales::percent) +
  labs(
    y = "Cumlative percentage", 
    x = "Food density",
    title = "Cumulative distribution of food density"
  )

p1 + p2

```

#### Meal interval

```{r distr_meal_interval, fig.height = 4, fig.width = 8}
analytic_df %>% 
  filter(!is.na(meal_interval_hr)) %>% 
  summarize(
    min     = min(meal_interval_hr),
    Q1      = quantile(meal_interval_hr, prob = 0.25),
    Median  = quantile(meal_interval_hr, prob = 0.50),
    Mean    = mean(meal_interval_hr),
    SD      = sd(meal_interval_hr),
    Q3      = quantile(meal_interval_hr, prob = 0.75),
    max     = max(meal_interval_hr)
  ) %>% 
  knitr::kable(col.names = c("Min", "Q1", "Median", "Mean", "SD", "Q3", "Max"), digits = 2)

# Histogram
p1 <- analytic_df %>% 
  filter(!is.na(meal_interval_hr)) %>% 
  ggplot(aes(x = meal_interval_hr)) +
  geom_histogram(bins = 25) +
  labs(
    y = "Number of meals", 
    x = "Meal interval (hours)",
    title = "Histogram of meal interval"
  )

# ECDF
p2 <- analytic_df %>% 
  filter(!is.na(meal_interval_hr)) %>% 
  ggplot(aes(x = meal_interval_hr)) +
  stat_ecdf() +
  scale_y_continuous(labels = scales::percent) +
  labs(
    y = "Cumlative percentage", 
    x = "Meal interval (hours)",
    title = "Cumulative distribution of meal interval"
  )

p1 + p2
```

### Avocado intake

#### Distribution of avocado intake

* A histogram of avocado intake (grams) per meal/snack is shown below
  * The distribution of avocado intake per meal was trimodal, with the three largest bars corresponding to:
    * No avocado (0 grams) -- 70% of all meals
    * One-half (84 grams) -- 7.4%
    * One whole avocado (168 grams) -- 20%
  * The max avocado intake per meal was 504 grams

```{r avocado_histogram}

# Histogram of avocado intake
analytic_df %>% 
  ggplot(aes(x = avocado_gram)) +
  geom_histogram() +
  labs(
    x = "Avocado consumption (grams/meal)", 
    y = "Frequency",
    title = "Histogram of avocado consumption"
  )

```

* Given the distribution of avocado, meals/snaks were categorized into the following 3 groups by avocado consumption:

```{r avocado_meal_freq}

# Frequency of meals with avocado 
# Avocado intake categories
analytic_df %>% 
  count(avocado_intake_cat) %>% 
  mutate(percent = n / sum(n) * 100) %>% 
  knitr::kable(digits = 1, col.names = c("Avocado intake", "n", "%"))

```

#### Food kcal by meal type and avocado intake

* Mean (SD) food kcal by meal type and avocado intake are shown below:
  * Food kcal increases with avocado intake for both meals and snacks

```{r food_kcal_by_mealtype_avocado}

tab4_2 <- analytic_df %>%
  group_by(meal_or_snack, meal_type, avocado_intake_cat) %>%
  summarise(
    n    = n(),
    mean = mean(food_kcal, na.rm = TRUE),
    sd   = sd(food_kcal,   na.rm = TRUE),
    .groups = "drop"
  )


tab4_2 %>%  
  mutate(
    avocado_label = case_when(
      avocado_intake_cat == "No avocado at all"        ~ "EO with no avocado", 
      avocado_intake_cat == "<1 avocado"               ~ "EO with <1 avocado", 
      avocado_intake_cat == "1 whole avocado or more"  ~ "EO with 1+ avocado"
    ),
    mean_sd = paste0(round(mean), " ± ", round(sd))
  ) %>%
  select(meal_type, avocado_label, n, mean_sd) %>%
  pivot_wider(
    names_from  = avocado_label,
    values_from = c(n, mean_sd)
  ) %>%
  select(
    `Meal type`               = meal_type,
    `No avocado:<br> N`           = `n_EO with no avocado`,
    `No avocado:<br> Mean ± SD`   = `mean_sd_EO with no avocado`,
    `<1 avocado:<br> N`          = `n_EO with <1 avocado`,
    `<1 avocado:<br> Mean ± SD`  = `mean_sd_EO with <1 avocado`,
    `1+ avocado:<br> N`          = `n_EO with 1+ avocado`,
    `1+ avocado:<br> Mean ± SD`  = `mean_sd_EO with 1+ avocado`
  ) %>% 
  knitr::kable()

# Figure
tab4_2 %>% 
  ggplot(aes(x = meal_type, y = mean, fill = avocado_intake_cat)) +
  geom_bar(stat = "identity", position = "dodge", width = 0.7) +
  scale_y_continuous(limits = c(0, 1000), breaks = 1:5 * 200) +
  theme(legend.position = "bottom") +
  labs(x = "", y = "Mean food kcal per meal (kcal)", fill = "") +
  facet_wrap(~ meal_or_snack, scales = "free_x")

```

#### Food density by meal type and avocado intake

* Mean (SD) food density by meal type and avocado intake are shown below:
  * Food density decreases with avocado intake for meals

```{r food_density_by_mealtype_avocado}

tab4_2_fd <- analytic_df %>%
  group_by(meal_or_snack, meal_type, avocado_intake_cat) %>%
  summarise(
    n    = n(),
    mean = mean(food_density, na.rm = TRUE),
    sd   = sd(food_density,   na.rm = TRUE),
    .groups = "drop"
  )


tab4_2_fd %>%  
  mutate(
    avocado_label = case_when(
      avocado_intake_cat == "No avocado at all"        ~ "EO with no avocado", 
      avocado_intake_cat == "<1 avocado"               ~ "EO with <1 avocado", 
      avocado_intake_cat == "1 whole avocado or more"  ~ "EO with 1+ avocado"
    ),
    mean_sd = paste0(round(mean, 2), " ± ", round(sd, 2))
  ) %>%
  select(meal_type, avocado_label, n, mean_sd) %>%
  pivot_wider(
    names_from  = avocado_label,
    values_from = c(n, mean_sd)
  ) %>%
  select(
    `Meal type`               = meal_type,
    `No avocado:<br> N`           = `n_EO with no avocado`,
    `No avocado:<br> Mean ± SD`   = `mean_sd_EO with no avocado`,
    `<1 avocado:<br> N`          = `n_EO with <1 avocado`,
    `<1 avocado:<br> Mean ± SD`  = `mean_sd_EO with <1 avocado`,
    `1+ avocado:<br> N`          = `n_EO with 1+ avocado`,
    `1+ avocado:<br> Mean ± SD`  = `mean_sd_EO with 1+ avocado`
  ) %>% 
  knitr::kable()

# Figure
tab4_2_fd %>% 
  ggplot(aes(x = meal_type, y = mean, fill = avocado_intake_cat)) +
  geom_bar(stat = "identity", position = "dodge", width = 0.7) +
  theme(legend.position = "bottom") +
  labs(x = "", y = "Mean food density (kcal/g)", fill = "") +
  facet_wrap(~ meal_or_snack, scales = "free_x")

```

#### Prior avocado intake

* Among 3,694 meals/snacks (excluding the first eating occasions of the day), 69% of the meals/snack did not have any avocado intake at the preceding eating occasion
  * The meal interval appears to increase with greater avocado intake at the preceding eating occasion
  * The mean energy intake (from food) was lowest when less than one avocado was consumed at the preceding eating occasion

```{r prior_avocado_meal_interval_food_kcal}

# Prior avocado intake and its effect on meal interval and food kcal 
analytic_df %>% 
  filter(!is.na(lag_avocado_intake_cat)) %>%
  group_by(lag_avocado_intake_cat) %>% 
  summarize(
    n = n(),
    pct = n / nrow(.) * 100,
    mean_meal_interval_hr = mean(meal_interval_hr, na.rm = TRUE),
    sd_meal_interval_hr   =  sd(meal_interval_hr, na.rm = TRUE),
    mean_food_kcal        = mean(food_kcal, na.rm = TRUE),
    sd_food_kcal          =  sd(food_kcal, na.rm = TRUE)
  ) %>% 
  mutate(
    mean_sd_meal_interval = paste0(sprintf("%.2f", mean_meal_interval_hr), " ± ", sprintf("%.2f", sd_meal_interval_hr)),
    mean_sd_food_kcal     = paste0(sprintf("%.1f", mean_food_kcal), " ± ", sprintf("%.1f", sd_food_kcal))
  ) %>% 
  select(lag_avocado_intake_cat, n, pct, starts_with("mean_sd")) %>% 
  knitr::kable(digits = 2, col.names = c("Prior avocado intake", "n", "%", "Mean meal interval (SD)", "Mean food kcal (SD)"))

```

* When stratified by meal type, the meal interval increased with greater avocado intake at the preceding eating occasion for midday and evening meals, but not for morning meals

```{r prior_avocado_meal_interval}

tab <- analytic_df %>% 
  filter(!is.na(lag_avocado_intake_cat)) %>%
  group_by(meal_or_snack, meal_type, lag_avocado_intake_cat) %>% 
  summarize(
    n = n(),
    mean_meal_interval_hr = mean(meal_interval_hr, na.rm = TRUE),
    mean_food_kcal        = mean(food_kcal, na.rm = TRUE),
  )

tab %>% 
  ggplot(aes(x = meal_type, y = mean_meal_interval_hr, fill = lag_avocado_intake_cat)) +
  geom_bar(stat = "identity", position = "dodge") +
  facet_wrap(~meal_or_snack, scales = "free_x") +
  theme(legend.position = "bottom") +
  labs(x = "", y = "Mean meal interval (hour)", fill = "Prior avocado intake")

```

* When stratified by meal type, the mean energy content was lowest when less than one avocado was consumed at the preceding eating occasion for all meals (morning, midday, and evening meals)
  * For snacks, the mean energy content was highest when no avocado was consumed at the preceding eating occasion

```{r prior_avocado_food_kcal}

tab %>% 
  ggplot(aes(x = meal_type, y = mean_food_kcal, fill = lag_avocado_intake_cat)) +
  geom_bar(stat = "identity", position = "dodge") +
  facet_wrap(~meal_or_snack, scales = "free") +
  theme(legend.position = "bottom") +
  labs(x = "", y = "Mean energy content (kcal)", fill = "Prior vocado intake")

```


## Linear Mixed Models

### Descriptions of the models

* To examine the effect of avocado consumption on three outcomes -- meal energy intake (kcal from food), dietary energy density (kcal/g), and mealtime interval (hours between consecutive eating occasions) -- we fitted separate linear mixed effects models for each outcome using the `lme` function from the `nlme` package in R

* Each model accounted for the three-level nested structure of the data, where eating occasions (level 1) were nested within recall days (level 2), which were in turn nested within participants (level 3)

* The same covariance structure was applied across all three models, with an unstructured G matrix at the participant level to capture between-person variability in both baseline outcomes and recall-to-recall changes, and a continuous autoregressive structure for the R matrix to account for the correlation among meals consumed within the same recall day. This approach recognizes that two meals eaten close together in time are more likely to influence each other than two meals eaten far apart

### Covariates

* All models included the following covariates: 
  * Age at baseline (as continuous)
  * Sex (female as reference)
  * Race (Caucasian as reference)
  * BMI (as continuous)
  * Residential region (East/West coast, East coast as reference)
  * Day of the week (weekdays or weekends, weekday as reference)
  * Total daily kcal (as continuous, per 100 kcal)
  * Meal type (evening meal as reference)
  * Meal environment (4 levels as shown above; Home/Friend's home as reference)
  * Meal frequency (as continuous)
  
### Outcomes

* Three outcomes were examined: meal energy intake (kcal), dietary energy density (kcal/g), and mealtime interval (hours)
  * Given the severe right-skewed distribution of meal energy intake, this outcome was log-transformed prior to analysis
  * No transformation was applied to energy density or mealtime interval
  
* To examine the effect of avocado consumption on meal energy intake and dietary energy density, the models additionally included:
  * Avocado consumption at the same eating occasion as a categorical variable (no avocado as the reference category),
  * along with its interaction with meal type where statistically significant

* For the subsequent meal models:
  * Avocado intake at the preceding eating occasion was included as the main exposure,
  * again with its interaction with meal type retained where statistically significant

* For each outcome, estimated marginal means (EMMs) with 95% confidence intervals were derived from the fitted models using the `emmeans` package in R, averaging over all other covariates in the model at their observed distributions
  * For meal energy intake, where a log transformation was applied prior to analysis, EMMs were back-transformed to the original scale for interpretability and are therefore reported as geometric means with 95% confidence intervals

* The assumptions of the linear mixed models were examined using normalized (decorrelated) residuals. No severe departures from normality were observed. All analyses were conducted in R version 4.5.3.

## Model results

### Effect of avocado consumption on meal energy intake

* The interaction between meal type and avocado intake was highly significant (p <.0001),
  * suggesting that the effects of avocado intake on meal energy content differed across meal types

* Among the covariates: 
  * Age, sex, race/ethinicity, BMI, region, and day of the week were not statistically significant
  * Meal environment, meal frequency, daily energy intake were all highly significant (p <.0001)
    * Meal environment: Significantly higher kcal for "Deli/Take-out/Restaurant" and "Daycare/Party/Travelling/Other"
    * Meal frequency: Higher the frequency, lower the energy intake
    * Daily energy intake: Higher the daily kcal, higher the meal kcal

* The table below presents estimated marginal means of meal energy intake by meal type and avocado intake category, 
  * adjusted for all covariates in the model and 
  * back-transformed to the original scale (kcal)
  
* The last two columns show Tukey-adjusted p-values comparing against the no-avocado meals and against the <1 avocado meals, respectively
  * Across all meal types, meal energy content increased with avocado intake
  * Most pairwise comparisons within the same meal type were significant, with two exceptions:
    * <1 avocado vs 1+ avocado comparison in the morning meal
    * <1 avocado vs 1+ avocado comparison in the late-night snack


```{r lmm_avocado_log_kcal}

# Mixed models ------------------------------------------------------------

## Food kcal --------------------------------------------------------------

# Data for model
analytic_df2 <- analytic_df %>% 
  arrange(pid, visitcode, cmealid) %>% 
  filter(!is.na(meal_place)) %>% 
  mutate(
    log_food_kcal = log(food_kcal),
    meal_type = relevel(meal_type, ref = "Evening meal"),
    cmealid = factor(cmealid),
    daily_kcal100 = kcal_per_day / 100,
    meal_time_hr = tmtime_fixed / 3600,
    meal_period = case_when(
      meal_type %in% c("Morning meal", "Morning snack") ~ 1,
      meal_type %in% c("Midday meal", "Midday snack") ~ 2,
      meal_type %in% c("Evening meal", "Late night snack") ~ 0
    ),
    meal_period = factor(meal_period, labels = c("Evening", "Morning", "Midday"))
    )

# Covariates
covars <- c(
  "age", 
  "sex", 
  "race", 
  "bmi", 
  "region", 
  "weekday", 
  "meal_place", 
  "meal_freq", 
  "daily_kcal100"
)

# Fit the model with covariates only
# Create formula
fm <- formula(paste("log_food_kcal ~", paste(covars, collapse = " + ")))

# Run covariates-only model
log_food_kcal_0 <- lme(
  fixed       = fm,
  random      = ~ 1 + visitcode | pid,          
  correlation = corCAR1(form = ~ meal_time_hr | pid/visitcode),
  data        = analytic_df2,
  method      = "REML",
  control = lmeControl(
    opt = "optim",
    maxIter = 200,
    msMaxIter = 200
  )
)

# Add avocado intake * meal type interaction
log_food_kcal_1 <- update(log_food_kcal_0, . ~ . + avocado_intake_cat * meal_type)

# Estimated marginal means
emmeans_avocado <- emmeans(log_food_kcal_1, ~avocado_intake_cat | meal_type, weights = "proportional")
pairs_avocado   <- pairs(emmeans_avocado, adjust = "tukey")
emmeans_tab     <- make_emmeans_tab(emmeans_avocado, pairs_avocado, exponentiate = TRUE, digits = 1)

# Table format
emmeans_tab %>%
  select(-meal_or_snack) %>% 
  select(meal_type, everything()) %>%
  mutate(across(where(is.character), ~replace_na(., ""))) %>%
  kable(
    col.names = c(
      "Meal type", 
      "Avocado intake", 
      "Adjusted mean", 
      "Lower 95% CL", 
      "Upper 95% CL", 
      "p-val: vs No avocado",
      "p-val: vs <1 avocado"
      ),
    align     = c("l", "l", "r", "r", "r", "r", "r")
  ) %>%
  kable_styling() %>%
  collapse_rows(columns = 1, valign = "top")

```

```{r lmm_avocado_log_kcal_plot}

# Bar chart
emmeans_tab %>% 
  ggplot(aes(x = meal_type, y = emmean, fill = avocado_intake_cat)) +
  geom_col(position = "dodge") +
  geom_errorbar(
    aes(ymin = lower.CL, ymax = upper.CL),
    width = 0.2,
    position = position_dodge(0.9)
  ) +
  facet_wrap(~ meal_or_snack, scales = "free_x") +
  labs(
    title = "Estimated marginal means (with 95% CI) of food kcal",
    x = "", 
    y = "Adjusted mean energy content (kcal/meal)",
    fill = "") +
  theme(legend.position = "bottom")

```

### Effect of avocado consumption on energy density

* The interaction between meal type and avocado intake was highly significant (p <.0001),
  * suggesting that the effects of avocado intake on energy density differed across meal types

* Among the covariates: 
  * Age, sex, BMI, day of the week, and daily were not statistically significant
  * The following covariates were significant:
    * Race/ethnicity   (p = 0.038, lower density among Asians) 
    * Region           (p = 0.021, higher density in the west coast) 
    * Meal environment (p = 0.004, higher density in "Daycare/Party/Travelling/Other")
    * Meal frequency   (p = 0.002, higher the frequency, lower the density)

* The table below presents estimated marginal means of energy density by meal type and avocado intake category, adjusted for all covariates in the model
  
* Overall, energy density tended to be highest in meals that did not include any avocado 
  * Among snacks, energy density was significantly higher in snacks containing no avocado compared to snacks containing <1 avocado or 1+ avocados, across all snack periods (morning, midday, and late-night). The difference between the latter two was not statistically significant.
  * Among meals, significant differences were found in three comparisons:
    * no avocado vs <1 avocado in the morning meal
    * no avocado vs 1+ avocado in the morning meal
    * <1 avocado vs 1+ avocado in the midday meal

```{r lmm_avocado_energy_density}

## Energy density ---------------------------------------------------------

# Run covariates-only model
energy_density_0 <- update(
  log_food_kcal_0, 
  food_density ~ ., 
  data = analytic_df2,
  subset = !is.na(food_density)
)

# Add avocado intake * meal type interaction
energy_density_1 <- update(energy_density_0, . ~ . + avocado_intake_cat * meal_type)

# Estimated marginal means
emmeans_avocado <- emmeans(energy_density_1, ~avocado_intake_cat | meal_type, weights = "proportional")
pairs_avocado   <- pairs(emmeans_avocado, adjust = "tukey")
emmeans_tab     <- make_emmeans_tab(emmeans_avocado, pairs_avocado, exponentiate = FALSE, digits = 2)

# Table format
emmeans_tab %>%
  select(-meal_or_snack) %>% 
  select(meal_type, everything()) %>%
  mutate(across(where(is.character), ~replace_na(., ""))) %>%
  kable(
    col.names = c(
      "Meal type", 
      "Avocado intake", 
      "Adjusted mean", 
      "Lower 95% CL", 
      "Upper 95% CL", 
      "p-val: vs No avocado",
      "p-val: vs <1 avocado"
      ),
    align     = c("l", "l", "r", "r", "r", "r", "r")
  ) %>%
  kable_styling() %>%
  collapse_rows(columns = 1, valign = "top")

```

```{r lmm_avocado_energy_density_plot}

# Bar chart
emmeans_tab %>% 
  ggplot(aes(x = meal_type, y = emmean, fill = avocado_intake_cat)) +
  geom_col(position = "dodge") +
  geom_errorbar(
    aes(ymin = lower.CL, ymax = upper.CL),
    width = 0.2,
    position = position_dodge(0.9)
  ) +
  facet_wrap(~ meal_or_snack, scales = "free_x") +
  labs(
    title = "Estimated marginal means (with 95% CI) of energy density",
    x = "", 
    y = "Adjusted mean energy density (kcal/gram)",
    fill = "") +
  theme(legend.position = "bottom")

```


### Effect of preceding avocado consumption on meal energy intake

* The interaction between meal type and preceding avocado intake was statistically significant (p = 0.010),
  * suggesting that the effects of preceding avocado intake on meal energy content differed across meal types

* Among the covariates: 
  * Age, sex, race/ethinicity, BMI, region, and day of the week were not statistically significant
  * Meal environment, meal frequency, daily energy intake were all highly significant (p <.0001)
    * Meal environment: Significantly higher kcal for "Deli/Take-out/Restaurant"
    * Meal frequency: Higher the frequency, lower the energy intake
    * Daily energy intake: Higher the daily kcal, higher the meal kcal

* The table below presents estimated marginal means of meal energy intake by meal type and preceding avocado intake category, 
  * adjusted for all covariates in the model and 
  * back-transformed to the original scale (kcal)
  
* For midday meals, the absence of avocado in the preceding eating occasion was associated with significantly higher energy content in the subsequent meal, compared to when any amount of avocado was consumed
  * No significant differences were found for the morning or evening meals
  
* For morning snacks, a similar pattern was observed -- the absence of avocado in the preceding eating occasion was associated with significantly higher energy content in the subsequent eating occasion
  * For midday snacks, energy content was significantly lower when 1+ avocado was consumed in the preceding eating occasion compared to when none or <1 avocado was consumed
  

```{r lmm_lag_avocado_log_kcal}

# Effect of avocado intake on the subsequent meal -------------------------

## Food kcal --------------------------------------------------------------

# Run covariates-only model
log_food_kcal_0 <- lme(
  fixed       = fm,
  random      = ~ 1 + visitcode | pid,          
  correlation = corCAR1(form = ~ meal_time_hr | pid/visitcode),
  data        = analytic_df2,
  subset      = !is.na(lag_avocado_intake_cat),
  method      = "REML",
  control = lmeControl(
    opt = "optim",
    maxIter = 200,
    msMaxIter = 200
  )
)

# Add avocado intake * meal type interaction
log_food_kcal_1 <- update(log_food_kcal_0, . ~ . + lag_avocado_intake_cat * meal_type)

# Estimated marginal means
emmeans_avocado <- emmeans(log_food_kcal_1, ~lag_avocado_intake_cat | meal_type, weights = "proportional")
pairs_avocado   <- pairs(emmeans_avocado, adjust = "tukey")
emmeans_tab     <- make_emmeans_tab(emmeans_avocado, pairs_avocado, exponentiate = TRUE, digits = 1)

# Table format
emmeans_tab %>%
  select(-meal_or_snack) %>% 
  select(meal_type, everything()) %>%
  mutate(across(where(is.character), ~replace_na(., ""))) %>%
  kable(
    col.names = c(
      "Meal type", 
      "Preceding avocado intake", 
      "Adjusted mean", 
      "Lower 95% CL", 
      "Upper 95% CL", 
      "p-val: vs No avocado",
      "p-val: vs <1 avocado"
      ),
    align     = c("l", "l", "r", "r", "r", "r", "r")
  ) %>%
  kable_styling() %>%
  collapse_rows(columns = 1, valign = "top")

```

```{r lmm_lag_avocado_log_kcal_plot}

# Bar chart
emmeans_tab %>% 
  ggplot(aes(x = meal_type, y = emmean, fill = lag_avocado_intake_cat)) +
  geom_col(position = "dodge") +
  geom_errorbar(
    aes(ymin = lower.CL, ymax = upper.CL),
    width = 0.2,
    position = position_dodge(0.9)
  ) +
  facet_wrap(~ meal_or_snack, scales = "free_x") +
  labs(
    title = "Estimated marginal means (with 95% CI) of energy density",
    x = "", 
    y = "Adjusted mean energy density (kcal/gram)",
    fill = "Preceding EO") +
  theme(legend.position = "bottom")

```


### Effect of preceding avocado consumption on energy density

* The interaction between meal type and preceding avocado intake was not significant (p = 0.09),

* Among the covariates: 
  * Age, sex, race/ethinicity, BMI, region, day of the week, and meal frequency were not statistically significant
  * Meal environment (p = 0.0002) and daily energy intake (p <.0001) were highly significant 
    * Meal environment: Significantly higher density for "Daycare/Party/Travelling/Other"
    * Daily energy intake: Energy density was positively associated with daily energy intake

* The table below presents estimated marginal means of energy density by preceding avocado intake, adjusted for all covariates in the model
  * Energy density tended to increase with greater avocado consumption, with a significant difference observed between meals containing no avocado and meals containing 1+ avocados in the preceding eating occasion (p = 0.012).
  
```{r lmm_lag_avocado_energy_density}

# Run covariates-only model
energy_density_0 <- update(
  log_food_kcal_0, 
  food_density ~ ., 
  subset = !is.na(lag_avocado_intake_cat) & !is.na(food_density))

# Add avocado intake * meal type interaction
energy_density_1 <- update(energy_density_0, . ~ . + lag_avocado_intake_cat * meal_type)

# Remove avocado intake * meal type interaction
energy_density_2 <- update(energy_density_1, . ~ . - lag_avocado_intake_cat:meal_type)

# Estimated marginal means
emmeans_avocado <- emmeans(energy_density_2, ~lag_avocado_intake_cat, weights = "proportional")

pairs_df <- pairs(emmeans_avocado, adjust = "tukey") %>%
  summary() %>%
  as.data.frame() %>%
  select(contrast, p.value) %>%
  mutate(p.value = format.pval(round(p.value, 4), digits = 4, eps = 0.0001, scientific = FALSE)) %>%
  pivot_wider(names_from = contrast, values_from = p.value) %>%
  rename(
    vs_no_avoc  = 1,  # "No avocado at all - <1 avocado"
    vs_no_avoc2 = 2,  # "No avocado at all - 1 whole avocado or more"
    vs_lt_1avoc = 3   # "<1 avocado - 1 whole avocado or more"
  )

emmeans_tab_overall <- summary(emmeans_avocado) %>%
  as.data.frame() %>%
  select(lag_avocado_intake_cat, emmean, lower.CL, upper.CL) %>%
  mutate(across(c(emmean, lower.CL, upper.CL), ~ round(., 2))) %>%
  mutate(
    vs_no_avoc = case_when(
      lag_avocado_intake_cat == levels(lag_avocado_intake_cat)[2] ~ pairs_df$vs_no_avoc,
      lag_avocado_intake_cat == levels(lag_avocado_intake_cat)[3] ~ pairs_df$vs_no_avoc2,
      TRUE ~ NA_character_
    ),
    vs_lt_1avoc = case_when(
      lag_avocado_intake_cat == levels(lag_avocado_intake_cat)[3] ~ pairs_df$vs_lt_1avoc,
      TRUE ~ NA_character_
    )
  )

emmeans_tab_overall %>% 
  mutate(across(where(is.character), ~replace_na(., ""))) %>%
  kable(
    col.names = c(
      "Preceding avocado intake", 
      "Adjusted mean", 
      "Lower 95% CL", 
      "Upper 95% CL", 
      "p-val: vs No avocado",
      "p-val: vs <1 avocado"
      ),
    align     = c("l", "r", "r", "r", "r", "r")
  ) %>%
  kable_styling() %>%
  collapse_rows(columns = 1, valign = "top")

```


### Effect of preceding avocado consumption on mealtime interval

* The interaction between meal type and preceding avocado intake was highly significant (p = 0.0003),
  * suggesting that the effects of preceding avocado intake on meal interval differed across meal types

* Among the covariates: 
  * Age, race/ethinicity, BMI, region, day of the week, meal environment, and daily kcal were not significant
  * Sex and meal frequenc were both highly significant (p <.0001)
    * Sex: Males had longer meal intervals
    * Meal frequency: Higher the frequency, shorter the interval

* The table below presents estimated marginal means of meal interval (hours) by meal type and preceding avocado intake category, adjusted for all covariates in the model 

* Meals:
  * For morning meals, there were no significant differences in meal interval by preceding avocado intake
  * For midday meals, prior consumption of 1+ avocados was associated with a significantly longer mealtime interval compared to no prior avocado consumption
  * Among evening meals, prior consumption of any amouhnt of avocado was associated with a significantly longer mealtime interval compared to no prior avocado consumption. No significant differences in interval was observed between <1 avocado and 1+ avocados at the preceding eating occasion

* For snacks, no significant differences were observed across any period, including morning, midday, and late-night snacks
  
```{r lmm_lag_avocado_meal_interval}

# Run covariates-only model
meal_interval_0 <- update(
  log_food_kcal_0, 
  meal_interval_hr ~ ., 
  subset = !is.na(lag_avocado_intake_cat) & !is.na(meal_interval_hr))

# Add avocado intake * meal type interaction
meal_interval_1 <- update(meal_interval_0, . ~ . + lag_avocado_intake_cat * meal_type)

# Estimated marginal means
emmeans_avocado <- emmeans(meal_interval_1, ~lag_avocado_intake_cat | meal_type, weights = "proportional")
pairs_avocado   <- pairs(emmeans_avocado, adjust = "tukey")
emmeans_tab     <- make_emmeans_tab(emmeans_avocado, pairs_avocado, digits = 2)

# Table format
emmeans_tab %>%
  select(-meal_or_snack) %>% 
  select(meal_type, everything()) %>%
  mutate(across(where(is.character), ~replace_na(., ""))) %>%
  kable(
    col.names = c(
      "Meal type", 
      "Preceding avocado intake", 
      "Adjusted mean", 
      "Lower 95% CL", 
      "Upper 95% CL", 
      "p-val: vs No avocado",
      "p-val: vs <1 avocado"
      ),
    align     = c("l", "l", "r", "r", "r", "r", "r")
  ) %>%
  kable_styling() %>%
  collapse_rows(columns = 1, valign = "top")

```


```{r lmm_lag_avocado_meal_interval_plot}

# Bar chart
emmeans_tab %>% 
  ggplot(aes(x = meal_type, y = emmean, fill = lag_avocado_intake_cat)) +
  geom_col(position = "dodge") +
  geom_errorbar(
    aes(ymin = lower.CL, ymax = upper.CL),
    width = 0.2,
    position = position_dodge(0.9)
  ) +
  facet_wrap(~ meal_or_snack, scales = "free_x") +
  labs(
    title = "Estimated marginal means (with 95% CI) of meal interval (hour)",
    x = "", 
    y = "Adjusted mean meal interval (hour)",
    fill = "") +
  theme(legend.position = "bottom")

```

